---
title: "Literature Review"
author: "Chelsea Matthews"
date: "7th October 2019"
header-includes:
  \usepackage{float}
output: 
  pdf_document:
    number_sections: true
    toc: true
    toc_depth: 4
bibliography: bibliography.bib
---

\pagebreak

# Aim
To find evidence and provide guidance on the selection of sequencing technologies and assembly methods to provide adequate assembly quality within constrained resources. 
This will be achieved through a benchmarking study of de novo assembly pipelines using gold standard assembly tools and a range of sequencing platforms for the rice strain Minghui 63. This should include coverage requirements (and inferred associated sequencing costs), CPU hours for an assembly and a quality assessment of the resulting genome with a focus on the optimisation of available resources for a given project to meet time, cost and quality requirements where possible.   
Deliverables will include a range of reproducible pipelines for de-novo genome assembly and benchmarking assessment. 

# Introduction

There are two primary methods of genome assembly; a reference-guided assembly where reads are compared with a reference genome of the same species or a closely related organism, and a de novo assembly where overlapping reads are assembled from scratch [@kyriakidou_2018]. Reference-guided assembly (often termed resequencing when a reference genome of the same species is used) is computationally cheap compared with de novo assemblies which essentially require an 'all-to-all' comparison of reads. However, reference-guided assemblies are often plagued by reference bias, where variation not present in the reference genome is missed in the genome being assembled. De novo assembly requires a greater depth of genome coverage but solves the reference bias problem [@baker_2012; @chin_2019]. 

In bacteria, de novo assembly is fairly routine as the small genome size results in a low complexity of assembly which, in turn, results in a cheap assembly. As genome size increases, de novo assembly becomes much more expensive due to the increased complexity of the assembly and hence, the amount of work and time required. In a complex de novo assembly, the cost of the sequencing itself becomes insignificant (per kb) when compared with the cost of the bioinformaticians time. 

While a high quality assembly is obviously preferred, it is not always possible - or even necessary - depending on the biological question being investigated. Resource availability including money, knowledge and time, heavily influence what assembly quality can reasonably be achieved. There is a huge amount of research available regarding the differences between assembly tools and underlying algorithms [@cherukuri_2016; @giordano_2017; @khan_2018], the quality of genomes assembled using data from different sequencing platforms [@wenger_2019; @giordano_2017] and the error profiles of sequencing technologies [list of references here]. However, there is a distinct lack of information available to help a researcher maximise the quality of an assembly given their available resources so that their biological question can be fully addressed. In filling this gap, we review the current literature including current sequencing technologies, methods of measuring assembly quality, an investigation into different workflows and methods for assembly as well as ways to ensure workflow reproducibility. This information will then be used to design a benchmarking comparison of different methods for assembly including an analysis of sequencing costs, CPU hours, and the resulting assembly quality. 

# The First Human Assembly

The first human genome assembly was a mammoth task and began in 1990 with a budget of \$3 billion US [@HGP_2003]. The project completed two years ahead of schedule (and under budget!) in 2003 and resulted in an assembly that covered approximately 99% of the gene-containing regions of the genome and was 99.99% accurate [@HGP_2003]. This project was made possible by Sanger, Nickeln and Coulson when they published their gamechanging paper "DNA sequencing with chain-terminating inhibitors" in 1977. In this paper, they introduced what is now termed Sanger sequencing, a method of sequencing DNA that is relatively robust, easy to use and accurate and was the method of choice for the Human Genome Project. Sanger sequencing uses DNA Polymerase to synthesise the complementary strand of a single-stranded DNA molecule in the presence of a small number of chain-terminating dideoxynucleotides which results in a library of double stranded fragments terminating at various different locations. Through fractioning these fragments with electrophoresis, the resulting bands give the sequence of nucleotides (See Figure \ref{fig:sangerSeq} below for further detail on this process) [@sanger_1977; @frana_2002].

\ref{fig:sangerSeq}

```{r sangerSeq, fig.width=8,echo=FALSE,fig.cap="\\label{fig:sangerSeq}Sanger Sequencing technique.",out.extra='', fig.pos='H'}
library(png)
library(grid)
img <- readPNG("images/sanger_method.png")
 grid.raster(img)
```

**Source: @gauthier_2007.**

Sanger sequencing yields highly accurate reads (up to 99.999% accurate) [@shendure_2008] of approximately 800bp in length (up to a maximum of ~1000bp) [@sanger_length; @lutteropp_2017]. Errors tend to be primarily substitutions [@lutteropp_2017]. However, despite advances in sanger sequencing technology resulting in improved throughput and reduced cost, it was, and still is, prohibitively expensive (~\$7.50 AU per kbp for between 1 and 10 samples down to ~\$4.50 AU for between 97 and 480 samples (taken from @biobasic_price)) and slow. Enter high throughput sequencing. 

# High Throughput Sequencing

The next generation of sequencing technology began to emerge around the same time that large-scale chain-termination methods were being developed [@heather_2016] and was characterised by the ability to sequence millions of DNA fragments at the same time [@behjati_2013]. High throughput sequencing techniques can be subdivided into at least two sub-categories; second generation techniques and single molecule sequencing. 

## Second Generation

Illumina is by far the most commonly used second generation technology [@greenleaf_2014]. It provides a range of options for both benchtop and production scale sequencing and is capable of producing both single and paired end reads of various lengths (up to 250bp). Illumina uses a patented sequencing by synthesis technology. First, library preparation is completed by randomly fragmenting DNA and ligating adapters to both ends of each fragment. These fragments are then PCR amplified and purified to form the library. This library is then loaded onto a flow cell where each fragment is amplified into a cluster so that sequencing can begin. The sequencing process itself consists of imaging fluorescently tagged reversible terminators as they are incorporated and then cleaved from the fragment clusters by DNA polymerase [@buermans_2014]. See Figure \ref{fig:sbsSeqTech} below for more detail on cluster generation and sequencing. 

```{r sbsSeqTech,echo=FALSE,fig.cap="\\label{fig:sbsSeqTech}Illumina cluster generation and sequence imaging.",out.extra='', fig.pos='H'}
library(png)
library(grid)
img <- readPNG("images/illumina_method.png")
 grid.raster(img)
```

**Source: @illumina_image**

While this method results in high quality reads (~99.8 accuracy in forward reads and ~99.6 for reverse reads), the error distribution is not random. Substitutions are the main type of error, being more than 1000 times more likely than insertions or deletions and more likely to occur following the motifs 'CGG' and 'GGG' [@schirmer_2016]. Errors also tend to accumulate near the end of the read due to phasing issues during the run obscuring the true signal from a cluster. Many authors also report limitations due to preferential amplification where fragments with very low or very high GC content are amplified less well, shorter fragments are amplified preferentially and sequence mismatches in the priming sites of the fragments affects priming and hence, amplification [@aird_2011; @krehenwinkel_2017]. Sequencing by synthesis is cheap (costing between 2.4 and 2.9c/Mbp based on a recent quote from Illumina) and fast and it's introduction to the market made sequencing many times more affordable. So much so that in 2014, Illumina advertised the first \$1000 human genome [@illumina_1000genome] and costs have decreased yet further since then. However, due to the short length of illumina reads, repetitive areas of the genome are very difficult to assemble and Illumina only de novo assemblies are highly fragmented. 

## Single Molecule Sequencing

Single Molecule Sequencing was introduced to the market by Pacific Biosciences (PacBio) in 2010. This technique negated the need for amplification of DNA fragments and hence, removed the associated errors and biases [@schadt_2010]. Oxford Nanopore Technologies (ONT) also provides a single molecule sequencing platform (released in 2014) and both technologies remain in use today.  

### PacBio

PacBio sequencing uses a patented SMRT (Single-Molecule, Real-Time) cell and creates a 'movie' of the real-time incorporation of bases into a strand. To achieve this, a DNA polymerase is immobilised in a Zero Mode Waveguide and illuminated from below by a laser which enables the detection of the incorporation of individual phospholinked nucleotides into the strand against the bulk solution background [@eid_2009] (see Figure \ref{fig:pacBioSMRT} below for more detail).  

```{r pacBioSMRT, fig.width=8,echo=FALSE,fig.cap="\\label{fig:pacBioSMRT}SMRT DNA sequencing. Each SMRT cell consists of an aluminium plate covered with 'wells' and a glass bottom. A DNA polymerase is immobilised in the bottom of each well inside a Zero Mode Waveguide. The chained polymerase incorporates phospholinked nucleotides from the solution into the single stranded fragments of interest while being illuminated from below by a laser. Because the nucleotide incorporation is occuring inside of the Zero Mode Waveguide, the light emission from the excited phospholinked nucleotides can be detected through the glass on the bottom of the plate despite the saturation of phospholinked nucleotides in the background.",out.extra='', fig.pos='H'}
library(png)
library(grid)
img <- readPNG("images/pacbioSMRT.png")
 grid.raster(img)
```

**Source: @eid_2009**

In addition to the SMRT cell, PacBio also makes use of the SMRTbell template. The SMRTbell template (shown in Figure 4 below) allows the DNA polymerase to continue incorporating bases around a circular single stranded molecule containing both the sense and antisense strands. Using this method, a consensus read can be called from multiple passes of the molecule. A single-pass sequence from PacBio will have an accuracy of approximately 85% [@wei_2018] but by calling a consensus from multiple passes of the same fragment, very high quality reads are able to be generated. This is particularly successful due to the random nature of the error distribution in PacBio single pass reads. Errors are predominantly deletions and insertions but are randomly distributed [@eid_2009] and so are easily removed when a consensus sequence is generated [@travers_2010]. PacBio reads also don't display any GC bias and so can be used for uniform sequencing of an entire genome [@rhoads_2015]. 

#### Hifi Reads

PacBio cap produce two different types of reads; Hifi reads and Continuous Long Reads (CLRs). Hifi reads are generated by selecting insert fragments of up to 20kbp and calling a consensus sequence from the subreads generated (see Figure \ref{fig:hifi} below). The accuracy of a hifi read is dependent on the number of passes made by the DNA polymerase; the more passes, the greater the accuracy. For example, with an average of 10 passes over each fragment in a tightly distributed 20kbp library, the average hifi read accuracy was 99.8% [@wenger_2019] (see Figure \ref{fig:hifidist} below for the read distribution of this library). This is comparable to the accuracy of Illumina reads but with much greater length [@wenger_2019]. The length and quality of hifi reads means that they can be used to assemble high quality and highly contiguous de novo assemblies. 

```{r hifi,echo=FALSE,fig.cap="\\label{fig:hifi}Hifi Read generation. First, single stranded hairpin adapters are ligated to both ends of a double stranded fragment of DNA. Primers are then annealed and the DNA polymerase begins to extend the primer from one of the hairpin loops. The polymerase incorporates bases around the hairpin and displaces one strand of the double stranded fragment to form a circular single strand of DNA. The polymerase continues incorporating bases until it reaches the primer where it began. It displaces the primer and continues incorporating bases around and around the circular strand, sequencing both the sense and antisense strands multiple times. This allows a consensus to be called which allows reads of high quality to be generated.",out.extra='', fig.pos='H', fig.asp=0.5}

img <- readPNG("images/How-to-get-HiFi-reads_clear.png")
grid.raster(img)
```
**Source: @pacbio_hifi.**

The most recently released iteration of the PacBio sequencing platform is Sequel II. Based on personal communications, we could expect approximately 20Gbp of Hifi reads from a single 8M SMRT Cell on the Sequel II at a cost of ~$8000 AU per cell plus \$1100 AU for library prep which equates to ~\$0.46 AU per Mbp. 

```{r hifidist, fig.width=8,echo=FALSE,fig.cap="\\label{fig:hifidist}Hifi read length distribution (before consensus calling) from a 20kbp size selected library on the Sequel II System.",out.extra='', fig.pos='H'}
library(png)
library(grid)
img <- readPNG("images/hifi_readlengths.png")
 grid.raster(img)
```

**Source: @pacbio_smrt.**

#### Continuous Long-Reads

CLRs are generated when the insert fragments are particularly large (up to 100kbp). For example, a 35kb size selected library on the Sequel II System produced CLR reads with an N50 of over 50 kbp (see Figure \ref{fig:longreaddist} below for detail on read distribution) [@pacbio_smrt]. Because of their length, CLRs are particularly useful for de novo assembly (particularly of repetitive regions) and detection of structural variants. Because of the random error distribution in PacBio reads, a consensus across a number of long reads can also be generated to increase the accuracy of an assembly [@chin_2013]. 

In the absence of a quote from DNA link, the following was assumed. With an average throughout of ~100Gbp per 8M SMRT Cell [@dna_link] and assuming Genomic CLR Library Prep and Sequel II SMRT Cell 8M with 30-hour movie from the University of Washington (total cost of ~\$4230 US), sequencing is projected to cost approximately $0.06 AU per Mbp. However, this cost would likely be higher in reality.   

```{r longreaddist, fig.width=8,echo=FALSE,fig.cap="\\label{fig:longreaddist}PacBio read length distribution of a 35 kb size selected library.",out.extra='', fig.pos='H', fig.asp=0.8}
library(png)
library(grid)
img <- readPNG("images/pacbio_longread_lengthdist.png")
 grid.raster(img)
```

**Source: @pacbio_smrt**

### Oxford Nanopore Technologies

ONT is the other major provider of single molecule sequencing technology. The MinION device was released in 2014 [@mikheyev_2014] and costs just $1000 AU for a MinION starter pack containing a single flow cell. It reports data in real time beginning just two minutes after the sequencing run begins (with 10 minutes sample preparation using the Rapid Kit) and is literally pocket sized **(Reference Nanopore website here)**. ONT offers devices ranging from very small portable sequencers suitable for field work to benchtop sequencers such as PromethION which runs up to 48 flow cells at any one time. All ONT devices, regardless of size, work in the same way. A single strand of DNA is threaded through a nanopore and changes in current across the pore are measured (see Figure \ref{fig:nanopore} below for more details on nanopore sequencing). A basecalling algorithm is then used to determine the actual sequence of the fragment.

```{r nanopore, echo=FALSE,fig.cap="\\label{fig:nanopore}Nanopore sequencing. A) The double stranded fragment of DNA is unwound by the unwinding enzyme at the top of the nanopore so that a single strand of DNA is threaded through the nanopore by a molecular motor protein (not shown). Three bases primarily occupy the pore at any one time with some small influence from flanking nucleotides (Rang, Kloosterman, and Ridder 2018). As the strand moves through the pore, changes in current across the pore are measured and recorded. B) These changes in current are called a squiggle and require interpretation to determine the actual base sequence. This interpretation is done using a basecaller.",fig.asp=0.55}
library(png)
library(grid)
img <- readPNG("images/nanopore_method.png")
 grid.raster(img)
```

**Source: Adapted from @minion_comp_requirements and @gpfrich_2018**

There are two main sources of error in nanopore reads; errors which occur during sequencing and are inherent in the raw data and errors made by the basecalling algorithm [@rang_2018]. In the latter case, the necessary information is present in the raw data but is interpreted incorrectly [@rang_2018].

With current nanopore technology, the DNA strand doesn't move through the pore at a constant or predictable rate [@rang_2018] and so it can be difficult to determine whether two adjacent measurements are from the same or different k-mers. Current basecallers use recurrent neural networks for base prediction and require training to optimise parameters. Instances of basecall error can be reduced by training a basecaller on a closely related species [@wick_2019; @jain_2018] as genome structure, codon usage and DNA modifications can vary between species [@rang_2018]. 

#### 1D nanopore reads

By passing a single strand of DNA through a nanopore, a 1D read is obtained. Based on recent ONT technology, nanopore 1D reads have an accuracy of approximately 80% with estimates of mismatches between 45% and 60% of total errors [@weirather_2017; @bowden_2019]. Deletions were the next most common error followed by insertions. Errors in nanopore data also aren't randomly distributed. @weirather_2017 found that the mismatch TAG->TGG was the most prominent error followed by TAC->TGC while other mismatches were much less frequent. However, it should be noted that other authors have found that deletions are the most dominant error type with insertions remaining as the least common [@cretustancu_2017]. This author also found an increase in deletions in sequences overlapping homopolymers and that rates of deletions and mismatches increased in GC rich regions. Coverage was also lower in GC rich regions but was not as pronounced as with Illumina sequencing of the same genome. Throughput on a MinION device is ~15Gbp on average with a cost of between \$1335 and \$705 AU per flow cell depending on the quantity purchased [@ont_prices]. At \$705 per cell, this equates ~$0.046 AU per Mbp.   

#### 1D^2^ nanopore reads

ONT also offer 1D^2^ reads. This technique enables both the sense and anti-sense strands of a DNA fragment to be sequenced sequentially so that a consensus sequence can be generated [@patel_2018] (see Figure \ref{fig:1d2nan} below for a comparison of 1D and 1D^2^ sequencing methods). Using this method, base accuracy is increased to approximately 95% [@volden_2018]. However, not all fragments are sequenced twice using the 1D^2^ technique. On a single flow cell with 1D^2^ chemistry, @volden_2018 obtained 1 million 1D reads and 50,000 1D^2^ reads. While other authors report higher ratios of 1D^2^ to 1D reads, this demonstrates that the output of 1D^2^ reads can be quite low.
Working on the assumptions that 1D^2^ reads are simply 1D reads with two fold coverage and that 1D^2^ sequencing throughput is similar to 1D throughput, they would cost approximately \$0.092 per Mbp with all 1D reads produced costing ~$0.046 AU per Mbp.

```{r 1d2nan, fig.width=8,echo=FALSE,fig.cap="\\label{fig:1d2nan}1D vs $1D^2$ nanopore reads. For a 1D read, only the template strand (blue) is threaded through the nanopore (grey) by the molecular motor protein (green). For $1D^2$ reads, both the template and complement strand (orange) are sequenced by tethering the complement strand to the membrane.",out.extra='', fig.pos='H'}
library(png)
library(grid)
img <- readPNG("images/nanopore_1D2_edit.png")
 grid.raster(img)
```

**Source: Adapted from @delannoy_2017**

Base accuracy of nanopore reads is greatly improved using the 1D^2^ method but it is unable to match PacBio Hifi reads for accuracy (99.8%). With the current nanopore technology, it is not possible to directly sequence a circular fragment to generate more than two-fold coverage of a read while PacBio is able to sequence the fragment of interest many times when generating a Hifi read. 

# Measuring Assembly Quality

There are a massive number of projects currently being undertaken involving the de novo assembly of previously unsequenced organisms. Some of these include the Global Invertebrates Genomics Alliance (GIGA) [@gigacommunityofscientists_2014], the Genome 10K Project [@koepfli_2015] and, most recently, the Earth BioGenome Project with the lofty goal of sequencing 'all of Earth's eukaryotic biodiversity over a period of ten years' [@lewin_2018] which includes a de novo assembly of at least one member from each family. With such a large number of sequencing projects underway [@gigacommunityofscientists_2014; @lewin_2018; @koepfli_2015], it is essential that we are able to accurately measure the quality and completeness of a de novo assembly [@yang_2019].

## Contig N50

Probably the most commonly cited metric of assembly quality is contig N50 [@kyriakidou_2018] where a contig N50 of 1Mbp implies that contigs of 1Mbp and greater comprise 50% of the total assembly size. However, this can be misleading for a number of reasons. Firstly, because it is calculated with respect to the assembly size and not the genome size, it is inherently biased. If a large portion of the genome is missing from the assembly, the contig N50 will appear more favourable than it actually is. Essentially contig N50 only gives an indication of the distribution of the contigs that are present and the term indication is used loosely as Contig N50 can be easily manipulated. By using more or less stringent filtering of short contigs, contig N50 can be artificially increased or reduced, respectively. It can also be increased by specifying less stringent requirements for the concatenation of contigs [@gurevich_2013]. In the extreme case, concatenation of all contigs would result in an assembly peppered with misassemblies but with a very high Contig N50, implying that the assembly is particularly good. For these reasons, contig N50 alone isn't enough to measure assembly quality. 

## QUAST

QUAST, Quality Assessment Tool for genome assemblies improves on this by reporting contig Nx in addition to the total number of contigs, length of the largest contig, total length, GC content and the number of predicted genes in a de novo assembly (when the length of the genome is unknown) [@gurevich_2013]. QUAST also produces a number of plots including a cumulative length plot (see Figure \ref{fig:cumulative} below for an example). A cumulative length plot can be very useful in determining assembly quality, providing that the user has knowledge of the exact methods and processes used in the assembly process. While the additional information provided by QUAST does give a much better indication of the data distribution, it still doesn't give any information regarding base quality or the likely completeness of the assembly. 

```{r cumulative, fig.width=7, echo=FALSE, fig.cap="\\label{fig:cumulative}A Cumulative Length Plot gives a detailed breakdown of the distribution of assembly contigs and the height of the right most point indicates the total assembly size. Here, the blue line represents an assembly comprised primarily of longer contigs while the red line shows an assembly of the same length but comprised of much shorter contigs. ",out.extra='', fig.pos='H'}
library(ggplot2)

cumulative_length <- c(40,130,230,390,530,640,690,740,780,810,830,844,857,870,881,890,897,900,903,903)
contig_index <- c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20)
bad_contigs <- c(0,0,0,0,	0,3,10,30,50,80,140,210,290,390,505,655,765,845,880,900)
rice_contigs <- data.frame(cumulative_length, contig_index, bad_contigs)

ggplot(rice_contigs, aes(x=contig_index))+
  geom_line(aes(y=cumulative_length), color = "steelblue") +
  geom_line(aes(y=bad_contigs), color = "red") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) + 
  labs(x="Contig Index (contigs sorted largest to smallest)", y = "Cumulative Coverage (Mbp)", title = "Cumulative Length Plot") + theme_bw()

```

## BUSCO

Another approach to assessing assembly completeness is demonstrated by BUSCO (Benchmarking Universal Single-Copy Orthologs) which benchmarks the ortholog content of a new assembly against a curated ortholog set generated from a selection of organisms belonging to the same clade [@simo_2015; @waterhouse_2017]. BUSCO then reports the number of complete orthologs found with only a single copy, complete orthologs with multiple copies, fragmented orthologs and missing orthologs [@simo_2015]. The higher the number of complete single copy orthologs, the more complete the assembly is likely to be while high numbers of multiple copy orthologs are likely to indicate that the assembly of haplotypes is incorrect [@simo_2015]. Unlike the methods mentioned previously, this gives some indication of how complete an assembly is. 

- talk about haplotype problems ??

## LTR Retrotransposon content

Another method for assessing the completeness of an assembly (primarily in plants) is the LTR Assembly Index (LAI) [@ou_2018_lai]. Given that LRT Retrotransposons comprise a large portion of most plant genomes [@ou_2017] and that they are notoriously difficult to assemble due to their repetitive nature, a measure of assembly continuity of these regions is warranted [@ou_2018_lai]. Ou, Chen and Jiang noted that more complete assemblies had a higher proportion of intract LTR elements when compared with less complete genomes (2018).To leverage this fact, they used LTR_retriever [@ou_2018_retriever] to first identify all LTR elements, then identified all intact LTR elements and finally applied a transformation to estimate LAI (see Equation 1 below) where LAI values less than 10 indicate draft genomes, between 10 and 20 are reference quality and LAI scores over 20 indicate gold standard assemblies [@ou_2018_lai]. 

<br>

**Equation 1:**

$$LAI = \frac{Intact \space LTR \space element \space length}{Total \space LTR \space sequence \space length}\times 100 + 2.8138 \times (94 - Whole \space genome \space LTR \space identity)$$
<br>

As many new assemblies obtain very high BUSCO scores [@ou_2018_lai], this method helps to further differentiate between a good and a great assembly by evaluating the quality of one of the more difficult to assemble region of the genome, if only in plants. 

## Quality Metrics used by Major Sequencing Projects

It is also worth investigating the methods proposed for assessing assembly quality in some of the major sequencing projects currently underway. The Earth BioGenome Project proposes to use a combination of contig N50, scaffold N50, a measure of chromosome assignment and a measure of basecall quality while avoiding assigning a direct classification (eg. gold, platinum, draft) [@lewin_2018]. GIGA proposes a similar approach but will also report, where possible, percent gaps, percent detection of conserved eukaryotic genes, a statistical assessment of assembly, alignment to any available syntenic or physical maps and mapping statistics of any availabla transcript data where possible [@gigacommunityofscientists_2014]. It appears as though both of these projects will be staying away from specific labelling of an assembly and will report only quantifiable measurements so that, as sequencing and assembly technology improves, these labels won't become redundant. 

# Generating an Assembly

## Sequencing Platform 

There are countless different paths to generating a de novo assembly but all of them begin with at least one type of sequencing data. In the beginning, the human genome was sequenced using primarily Sanger sequencing techniques [@lander_2001] but the advent of Second Generation techniques has made sequencing on a much larger scale possible due to higher throughput and reduced cost [@pollard_2018]. Sequencing by synthesis methods allowed a human genome to be sequenced in a day compared with years for the first human genome draft [@behjati_2013]. Illumina reads are accurate and well suited to sequencing non-repetitive regions of the genome but are less useful for the assembly of more complex and repetitive regions [@pollard_2018; @sovi_2016], particularly when repetitive regions exceed the length of the template fragment. They are not well suited to the discovery of structural variants [@cretustancu_2017]] and are further limited by preferential amplification of certain fragments. Single Molecule Sequencing was developed next and overcame some of these shortfalls while introducing new issues. Single Molecule Sequencing generates long reads which give context in extended repetitive regions and better enables the discovery of structural variants. Nanopore reads are cheap and long () but of low quality while PacBio is more expensive and offers both long (N50 of >50kbp), low quality reads, and long (up to 20kbp) high quality Hifi reads. Furthermore, the literature makes it very clear that the incorporation of at least some long reads into a de novo assembly greatly improves the final assembly quality when compared with using only short reads [@shin_2019; @jain_2018; @rhoads_2015]. 

## Assembly Algorithms

The two major classes of algorithms used in genome assembly are the Overlap-Layout-Consensus (OLC) and the de-Bruijn-Graph (dBG). The OLC algorithm was extensively used in the assembly of Sanger reads [@li_2012] while the dBG was relatively unknown in genome assembly until the development of second generation techniques where it really came into its own [@li_2012]. 

The OLC algorithm is relatively intuitive and is detailed below in Figure \ref{fig:olc}. It consists of three main steps: 

1) Find all overlaps between all reads
2) Layout all of the reads and overlaps on a graph 
3) Infer a consensus sequence from the graph 

```{r olc, echo=FALSE,fig.cap="\\label{fig:olc}The Overlap-Layout-Consensus algorithm. Reads are provided and the overlaps between reads are identified with an all-to-all pairwise alignment comparison. Then, these reads are graphed where each node represents a read and edges represent overlaps between reads. Then, the algorithm finds multiple hamiltonian paths (each vertex is visited exactly once) through the graph and takes the consensus of these paths which gives the final sequence.",out.extra='', fig.pos='H'}
library(png)
library(grid)
img <- readPNG("images/olcAlgorithm.png")
 grid.raster(img)
```

**Source: @commins_2009**

The dBG algorithm is not so intuitive as it begins with breaking reads up into even smaller pieces (see Figure \ref{fig:dbg} below). 

```{r dbg, echo=FALSE,fig.cap="\\label{fig:dbg}The deBruijn Graph method. Reads are cut up into k-mers and then split into a prefix and suffix including all but the last and first letter of the k-mer respectively. A de-Bruijn-Graph is then generated with all prefixes and suffixes as nodes, and edges between two nodes labelled with the k-mer having that particular prefix and suffix. Finally, an eulerian path (each node is visited exactly once) through the graph is found which indicates the final assembled sequence.",out.extra='', fig.pos='H'}
library(png)
library(grid)
img <- readPNG("images/debruijn.png")
 grid.raster(img)
```

**Source: @namiki_2012**

Figures \ref{fig:olc} and \ref{fig:dbg} above give a basic understanding of the algorithms underlying the OLC and dBG methods respectively but it is important to realise that there is additional complexity in implementations of these algorithms due to sequencing errors, repetitive regions, and sequencing heterozygous genomes or polyploids, among others. With respect to computational cost, the OLC algorithm is generally slower than the dBG due, in part, to the explicit all-to-all pairwise comparison of reads. The OLC algorithm is also slower due to the multiple sequence alignment required for generating the consensus sequence. This is a Hamiltonian path problem which is considered to be NP Complete (meaning that it can only be solved in polynomial time and hence, that the time required to solve this problem increases rapidly with the problem size) while the dBG algorithm uses an Eulerian path which is can be solved in linear time. **However, the OLC algorithm is better. **

## Assembly Methods

As mentioned earlier, the addition of at least some long reads to an assembly results in an increase in assembly quality and so only assembly methods meeting this criteria will be considered. 

When Single Molecule Sequencing was first developed, existing tools were unable to cope with the resulting very long, low quality reads. To get around this, researchers introduced algorithms to correct errors at the beginning of the workflow (polishing the long input reads) so that tools originally designed for Sanger Sequencing reads could be repurposed [@ye_2016]. Tools utilsing the OLC algorithm were of particular interest as the OLC method appears to result in higher quality assemblies [@cherukuri_2016] with higher coverage, larger average contig lengths and less contigs. This method of error correcting erraneous long reads before assembly is still commonly used [@goodwin_2015] with Illumina reads often being used for polishing errors out of Nanopore reads or PacBio CLRs. Other hybrid assembly methods include using long reads as scaffolds for the alignment of short reads [@ye_2016] and using short reads to polish errors from a long read assembly [@nanoporetech_2016]. 

It is obviously also possible to generate a de novo assembly from a single sequencing platform. PacBio Hifi reads, with their length and high quality, are well suited to this though nanopore technology provides higher throughput at ??. 

@weirather_2017 found that both nanopore and PacBio transcriptome analyses could benefit from a hybrid approach. 

## Resources Required to Generate a Genome Assembly

Figure \ref{fig:flowchart} below details the essential steps necessary for generating a de novo assembly as well as some of the costs associated with each step. Managing these costs while keeping the goal of a project in sight can be difficult 

```{r flowchart, echo=FALSE,fig.cap="\\label{fig:flow}Flowchart ",out.extra='', fig.pos='H',fig.asp=0.35}
library(png)
library(grid)
img <- readPNG("images/Flowchart.png")
 grid.raster(img)
```


# Workflow Reproducibility

Science depends on researchers being able to replicate the work of others but unfortunately, this isn't always possible. @baker_2016 found in a survey of 1500 scientists that 70% of researchers had been unable to reproduce another scientist's experiment at some point while, even more worrying, 50% were unable to reproduce one of their own experiments. In bioinformatics, reproducibility refers specifically to being able to obtain the exact same results for an analysis using the original dataset and code but, even when the data and code are supplied, this isn't always possible [@kim_2018]. @kim_2018 suggest that the main barrier to reproducibility is the incompatibility among environments, software versions and different languages while @kulkarni_2018 attribute these issues to the short half-life of bioinformatics software, complexity of workflows, uncontrolled effects induced by changes to system libraries and the imprecision in workflow descriptions. Workflow Management Systems seems to provide an answer to most of these issues. 

In general, workflow management systems provide infrastructure to help users build and execute workflows in a reproducible manner [@leipzig_2017]. They come in all different shapes and sizes with some common options including Snakemake, Nextflow and Galaxy [@perkel_2019]. Of the three, Galaxy is the only one with a Graphical User Interface and is a popular choice among researchers with limited programming knowledge [@giardine_2005]. Snakemake and Nextflow offer very similar features including containerisation (allows researchers to provide the exact version of software used in their analysis to ensure others can obtain the same results), tracking execution provenance, cloud compatibility and enabling parallel execution of tasks. They differ primarily in the way that the user visualises the workflow. Snakemake is rule-based and is similar to deciding what you want to eat and then working out how to make it. For example, if I want to serve lasagne for dinner, I first have to make the ragu; to make the ragu, I need to cook tomatos, meat and onion; and so on [@perkel_2019]. Nextflow uses a top-to-bottom model with a combination of channels and processes where channels pipe output from one process to the next [@ditommaso_2017]. Choosing between Snakemake and Nextflow essentially comes down to personal preference. They are both good choices for managing reproducibility, portability and scalability of a bioinformatics workflow [@perkel_2019].  

# Conclusion

Even when constrained to considering only the common sequencing platforms, there are major differences in read quality, error profiles, read lengths and cost. Illumina platforms generate high quality short reads at a very low cost while Nanopore reads are very long (up to 100kbp) and low quality but also at a relatively low price when compared with Sanger sequencing. PacBio offers two main sequencing options; very long, low quality reads (similar to nanopore) or shorter (~15kbp), very high quality reads (Hifi reads). From just these four sequencing types, there are a huge number of different ways that a de novo assembly can be generated including both single source and hybrid methods. Of particular interest are assemblies from PacBio Hifi reads, ONT reads, PacBio long reads, ONT reads with Illumina polishing and PacBio long reads with Illumina Polishing. Popular tools for these assemblies include Canu, Nanopolish, **(more here based on rest of review)**. In assessing the quality of a de novo assembly, there is no real consensus as to the best method but measures of contig length and distribution, scaffold length and distribution, conserved gene content and basecall quality are common measures. Workflow reproducibility is an essential part of any bioinformatics project as, without it, results aren’t able to be confirmed. When findings aren’t able to be ratified by other researchers, scientific progress is hampered. By using a workflow management system in an aware and responsible manner, a bioinformatician can ensure that, at the very least, their workflows and analysis are reproducible. 

```{python, engine.path = '/home/chelsea/anaconda3/bin/python3', echo=FALSE}
import sys
```

\pagebreak

# References
